---
title: "Week 4 - Homework"
author: "STAT 420, Summer 2020, D. Unger"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes  
  pdf_document: default
urlcolor: cyan
---


# Directions

Students are encouraged to work together on homework. However, sharing, copying or providing any part of a homework solution or code is an infraction of the University's rules on Academic Integrity. Any violation will be punished as severely as possible.

- Be sure to remove this section if you use this `.Rmd` file as a template.
- You may leave the questions in your final document.

***

## Exercise 1 (Using `lm`)

For this exercise we will use the data stored in [`nutrition-2018.csv`](nutrition-2018.csv). It contains the nutritional values per serving size for a large variety of foods as calculated by the USDA in 2018. It is a cleaned version totaling 5956 observations and is current as of April 2018.

The variables in the dataset are:

- `ID` 
- `Desc` - short description of food
- `Water` - in grams
- `Calories` 
- `Protein` - in grams
- `Fat` - in grams
- `Carbs` - carbohydrates, in grams
- `Fiber` - in grams
- `Sugar` - in grams
- `Calcium` - in milligrams
- `Potassium` - in milligrams
- `Sodium` - in milligrams
- `VitaminC` - vitamin C, in milligrams
- `Chol` - cholesterol, in milligrams
- `Portion` - description of standard serving size used in analysis

**(a)** Fit the following multiple linear regression model in `R`. Use `Calories` as the response and `Fat`, `Sugar`, and `Sodium` as predictors.

\[
Y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \beta_3 x_{i3} + \epsilon_i.
\]

Here,

- $Y_i$ is `Calories`.
- $x_{i1}$ is `Fat`.
- $x_{i2}$ is `Sugar`.
- $x_{i3}$ is `Sodium`.

Use an $F$-test to test the significance of the regression. Report the following:
 
- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.01$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

- The null and alternative hypotheses

$H_0: \beta_1 = \beta_2 = \beta_3 = 0$ vs $H_1 : \text{not } H_0$

- The value of the test statistic

```{r}
library(readr)
nutri = read.csv("nutrition-2018.csv")
```

```{r}
model = lm(Calories ~ Fat + Sugar + Sodium, data = nutri)
summary(model)
```
```{r}
# F-Statistic
summary(model)$fstatistic['value']
```

- The p-value of the test

```{r}
# P-value
library(broom)
summary(model)
glance(model)$p.value
```

P-value of the test is very close to 0, but is not zero. 

- A statistical decision at $\alpha = 0.01$


Since alpha is much larger than p-value, reject $H_0$.


- A conclusion in the context of the problem

There is signifcant amount of evidence that at least one predictor has *relationship* with the Calories(response) given that there are fat, sugar, and sodium in the model. 


**(b)** Output only the estimated regression coefficients. Interpret all $\hat{\beta}_j$ coefficients in the context of the problem.

```{r}
coef(model)[1]
```

This value is predicted calorie given that there are none of fat, sugar, and sodium in the model. 


```{r}
coef(model)[2]
```

An increase in fat per gram given that there are sugar and sodium in the model.


```{r}
coef(model)[3]
```

An increase in sugar per gram given that there are fat and sodium in the model.


```{r}
coef(model)[4]
```

An increase in Sodium per milligrams given that there are fat and sugar in the model .


**(c)** Use your model to predict the number of `Calories` in a Big Mac. According to [McDonald's publicized nutrition facts](https://www.mcdonalds.com/us/en-us/about-our-food/nutrition-calculator.html), the Big Mac contains 30g of fat, 9g of sugar, and 1010mg of sodium.

```{r}
predict(model, newdata = data.frame(Fat = 30, Sugar = 9, Sodium = 1010))
```


**(d)** Calculate the standard deviation, $s_y$, for the observed values in the Calories variable. Report the value of $s_e$ from your multiple regression model. Interpret both estimates in the context of this problem.

```{r}
s_y = sd(nutri$Calories)
s_y # Standard deviation of s_y
```
$s_y$ refers to the square root of variance of Calories. This shows the amount which Calories are dissemniated from the mean value.


```{r}
s_e = summary(model)$sigma
s_e # Residual standard error from multiple regression model
```
Residual standard error from multiple regression model of Calories equal to 80.8543. It means estimates of mean calories are “typically” off by 80.8543 *given that Sugar, Fat and Sodium are in the model*. 



**(e)** Report the value of $R^2$ for the model. Interpret its meaning in the context of the problem.

```{r}
summary(model)$r.squared
```

$R^2$ here means that the proportion of observed variation in calories that can be explained by the model equals to 0.7686281 *given that Sugar, Fat and Sodium are in the model*. 


**(f)** Calculate a 90% confidence interval for $\beta_2$. Give an interpretation of the interval in the context of the problem.

```{r}
confint(model, level=.9)['Sugar',]
```

We are $90%$ confident that for an increase in Sugar of 1 gram, the average increase in calories is between 3.783051 and 4.017983 *given that Fat and Sodium are in the model*.


**(g)** Calculate a 95% confidence interval for $\beta_0$. Give an interpretation of the interval in the context of the problem.

```{r}
confint(model, level=.95)['(Intercept)',]
```

We are $95\%$ confident that calories are between 97.69443 and 103.21768 where no sugar, no fat, no sodium are used *given that Sugar, Fat and Sodium are in the model*.


**(h)** Use a 99% confidence interval to estimate the mean Calorie content of a food with 23g of fat, 0g of sugar, and 400mg of sodium, which is true of a large order of McDonald's french fries. Interpret the interval in context.
 
```{r}
predict(model, newdata=data.frame(Fat = 23, Sugar = 0 , Sodium = 400), interval = "confidence", level=.99)
```
 
 We are $99 \%$ confident that calories are between $294.3532\%$ and $301.7224 \%$ where 23g of fat, 0g of sugar, and 400mg of sodium are used. 
 
 
**(i)** Use a 99% prediction interval to predict the Calorie content of a Crunchwrap Supreme, which has 21g of fat, 6g of sugar, and 1200mg of sodium according to [Taco Bell's publicized nutrition information](https://www.tacobell.com/nutrition/info). Interpret the interval in context.

```{r}
predict(model, newdata=data.frame(Fat = 21, Sugar = 6 , Sodium = 1200), interval = "prediction", level=.99)

```

We are $99 \%$ confident that calories are between $101.0345\%$ and $517.7786 \%$ where 21g of fat, 6g of sugar, and 1200mg of sodium are used. 


***

## Exercise 2 (More `lm` for Multiple Regression)

For this exercise we will use the data stored in [`goalies.csv`](goalies.csv). It contains career data for 462 players in the National Hockey League who played goaltender at some point up to and including the 2014-2015 season. The variables in the dataset are:
 
- `W` - Wins
- `GA` - Goals Against 
- `SA` - Shots Against 
- `SV` - Saves 
- `SV_PCT` - Save Percentage
- `GAA` - Goals Against Average
- `SO` - Shutouts 
- `MIN` - Minutes 
- `PIM` - Penalties in Minutes

For this exercise we will consider three models, each with Wins as the response. The predictors for these models are:

- Model 1: Goals Against, Saves
- Model 2: Goals Against, Saves, Shots Against, Minutes, Shutouts
- Model 3: All Available

```{r}
goalies = read_csv("goalies.csv")
model1 = lm(W ~ GA + SV, data=goalies)
model2 = lm(W ~ GA + SV + SA + MIN + SO, data=goalies)
model3 = lm(W ~ ., data=goalies)
```

**(a)** Use an $F$-test to compares Models 1 and 2. Report the following:


- The null hypothesis
$$H_0: \beta_{Shots Against} = \beta_{Minutes} = \beta_{Shutouts} = 0$$ 
- The value of the test statistic
```{r}
# anova(model1, model2)
anova(model1, model2)$F[2]
```

- The p-value of the test
```{r}
anova(model1, model2)
anova(model1, model2)$"Pr(>F)"[2]
```

P-value is a very small number

- A statistical decision at $\alpha = 0.05$

Since $\alpha >$ p-value, reject $H_0$.

- The model you prefer

I prefer model 2 because F statistic is very large while p-value is very small. Shots against, minutes, and shutouts seem to be related to wins.  


**(b)** Use an $F$-test to compare Model 3 to your preferred model from part **(a)**. Report the following:

- The null hypothesis
$$H_0: \beta_{\text{Save Percentage}} = \beta_{\text{Goals Against Average}} = \beta_{\text{Penalties in Minutes}} = 0$$

- The value of the test statistic
```{r}
anova(model2, model3)$"F"[2]
```

- The p-value of the test
```{r}
# anova(model2, model3)
anova(model2, model3)$"Pr(>F)"[2]
```

- A statistical decision at $\alpha = 0.05$

Since $\alpha > \text{p-value}$, reject $H_0$.

- The model you prefer

I prefer model3 to model2 because alpha is larger than p-value. 

**(c)** Use a $t$-test to test $H_0: \beta_{\texttt{SV}} = 0 \ \text{vs} \ H_1: \beta_{\texttt{SV}} \neq 0$ for the model you preferred in part **(b)**. Report the following:

- The value of the test statistic

```{r}
full = lm(W ~ GA + SA + SV + SV_PCT + GAA + SO + MIN + PIM, data=goalies)
coef(summary(full))['SV', 't value']
```

- The p-value of the test
```{r}
coef(summary(full))['SV', 'Pr(>|t|)']
```

- A statistical decision at $\alpha = 0.05$

Since $\alpha > \text{p-value}$, reject $H_0$.

***

## Exercise 3 (Regression without `lm`)

For this exercise we will once again use the `Ozone` data from the `mlbench` package. The goal of this exercise is to fit a model with `ozone` as the response and the remaining variables as predictors.

```{r}
data(Ozone, package = "mlbench")
Ozone = Ozone[, c(4, 6, 7, 8)]
colnames(Ozone) = c("ozone", "wind", "humidity", "temp")
Ozone = Ozone[complete.cases(Ozone), ]
```

**(a)** Obtain the estimated regression coefficients **without** the use of `lm()` or any other built-in functions for regression. That is, you should use only matrix operations. Store the results in a vector `beta_hat_no_lm`. To ensure this is a vector, you may need to use `as.vector()`. Return this vector as well as the results of `sum(beta_hat_no_lm ^ 2)`.

```{r}
X = cbind(1, Ozone$wind, Ozone$humidity, Ozone$temp)
y = cbind(Ozone$ozone)

beta_hat_no_lm = as.vector(solve(t(X) %*% X) %*% t(X) %*% y)

beta_hat_no_lm # Estimated beta parameters 
sum(beta_hat_no_lm^2) # Sum of squares of beta hat 
```

**(b)** Obtain the estimated regression coefficients **with** the use of `lm()`. Store the results in a vector `beta_hat_lm`. To ensure this is a vector, you may need to use `as.vector()`. Return this vector as well as the results of `sum(beta_hat_lm ^ 2)`.


```{r}
beta_hat_lm = as.vector(coef(lm(ozone ~ ., data=Ozone)))

beta_hat_lm 
sum(beta_hat_lm^2)
```


**(c)** Use the `all.equal()` function to verify that the results are the same. You may need to remove the names of one of the vectors. The `as.vector()` function will do this as a side effect, or you can directly use `unname()`.

```{r}
all.equal(beta_hat_lm, beta_hat_no_lm)
```


**(d)** Calculate $s_e$ without the use of `lm()`. That is, continue with your results from **(a)** and perform additional matrix operations to obtain the result. Output this result. Also, verify that this result is the same as the result obtained from `lm()`.

```{r}
y_hat = X %*% beta_hat_no_lm
sqrt(sum((y-y_hat)^2) / (nrow(Ozone)-length(beta_hat_no_lm)))
```


**(e)** Calculate $R^2$ without the use of `lm()`. That is, continue with your results from **(a)** and **(d)**, and perform additional operations to obtain the result. Output this result. Also, verify that this result is the same as the result obtained from `lm()`.

```{r}
SST = sum((y - mean(y))^2)
SSR = sum((y_hat - mean(y))^2)
R_Square = SSR/ SST
R_Square # R^2 calculated without lm

R_Square_lm = summary(lm(ozone ~ ., data=Ozone))$r.squared
R_Square_lm # R^2 calculated with lm

all.equal(R_Square, R_Square_lm) # Verification
```

Verified that obtained value without using lm equals the same with lm.

***

## Exercise 4 (Regression for Prediction)

For this exercise use the `Auto` dataset from the `ISLR` package. Use `?Auto` to learn about the dataset. The goal of this exercise is to find a model that is useful for **predicting** the response `mpg`. We remove the `name` variable as it is not useful for this analysis. (Also, this is an easier to load version of data from the textbook.)

```{r}
# load required package, remove "name" variable
library(ISLR)
Auto = subset(Auto, select = -c(name))
```

When evaluating a model for prediction, we often look at RMSE. However, if we both fit the model with all the data as well as evaluate RMSE using all the data, we're essentially cheating. We'd like to use RMSE as a measure of how well the model will predict on *unseen* data. If you haven't already noticed, the way we had been using RMSE resulted in RMSE decreasing as models became larger.

To correct for this, we will only use a portion of the data to fit the model, and then we will use leftover data to evaluate the model. We will call these datasets **train** (for fitting) and **test** (for evaluating). The definition of RMSE will stay the same

\[
\text{RMSE}(\text{model, data}) = \sqrt{\frac{1}{n} \sum_{i = 1}^{n}(y_i - \hat{y}_i)^2}
\]

where

- $y_i$ are the actual values of the response for the given data.
- $\hat{y}_i$ are the predicted values using the fitted model and the predictors from the data.

However, we will now evaluate it on both the **train** set and the **test** set separately. So each model you fit will have a **train** RMSE and a **test** RMSE. When calculating **test** RMSE, the predicted values will be found by predicting the response using the **test** data with the model fit using the **train** data. *__Test__ data should never be used to fit a model.*

- Train RMSE: Model fit with *train* data. Evaluate on **train** data.
- Test RMSE: Model fit with *train* data. Evaluate on **test** data.

Set a seed of `11`, and then split the `Auto` data into two datasets, one called `auto_trn` and one called `auto_tst`. The `auto_trn` data frame should contain 292 randomly chosen observations. The `auto_tst` data will contain the remaining observations. Hint: consider the following code:

```{r}
set.seed(11)
auto_trn_idx = sample(1:nrow(Auto), 292)
```

Fit a total of five models using the training data.

- One must use all possible predictors.
- One must use only `displacement` as a predictor.
- The remaining three you can pick to be anything you like. One of these should be the *best* of the five for predicting the response.

```{r}
names(Auto)
```
```{r}
auto_trn_idx = sort(auto_trn_idx)
auto_test_idx = rep(0, nrow(Auto) - length(auto_trn_idx))
n = 1
for (i in 1:nrow(Auto)) {
  if (is.element(i, auto_trn_idx) == FALSE) {
    auto_test_idx[n] = i
    n = n + 1
  }
}

model1 = lm(mpg ~ ., data=Auto[auto_trn_idx, ])
model2 = lm(mpg ~ displacement, data=Auto[auto_trn_idx, ])
model3 = lm(mpg ~ displacement+ weight + year + origin, data=Auto[auto_trn_idx, ])
model4 = lm(mpg ~ cylinders + displacement + horsepower + weight + acceleration + year, data=Auto[auto_trn_idx, ])
model5 = lm(mpg ~ cylinders + horsepower + weight + acceleration + year, data=Auto[auto_trn_idx, ])

```


For each model report the **train** and **test** RMSE. Arrange your results in a well-formatted markdown table. Argue that one of your models is the best for predicting the response.
```{r}
RMSE = function(criterion, model){
  y = Auto[criterion, ]$mpg
  y_hat = predict(model, newdata=data.frame(cylinders=Auto[criterion, ]$cylinders,
                                            displacement=Auto[criterion, ]$displacement,
                                            horsepower=Auto[criterion, ]$horsepower,
                                            weight=Auto[criterion, ]$weight,
                                            acceleration=Auto[criterion, ]$acceleration,
                                            year=Auto[criterion, ]$year,
                                            origin=Auto[criterion, ]$origin))
  sqrt(sum((y - y_hat)^2) / length(y))
}
```

```{r}
# Model1
# Predictor: All predictors 
trn_ml1 = RMSE(auto_trn_idx, model1)
tst_ml1 = RMSE(auto_test_idx, model1)

# Model2
# Predictor: Displacement 
trn_ml2 = RMSE(auto_trn_idx, model2)
tst_ml2 = RMSE(auto_test_idx, model2)

# Model3
# Predictor: displacement + weight + year + origin
trn_ml3 = RMSE(auto_trn_idx, model3)
tst_ml3 = RMSE(auto_test_idx, model3)

# Model4
# Predictor: cylinders + displacement + horsepower + weight + acceleration + year
trn_ml4 = RMSE(auto_trn_idx, model4)
tst_ml4 = RMSE(auto_test_idx, model4)

# Model5
# Predictor: cylinders + horsepower + weight + acceleration + year
trn_ml5 = RMSE(auto_trn_idx, model5)
tst_ml5 = RMSE(auto_test_idx, model5)

df = data.frame(Model1 = c(trn_ml1, tst_ml1),
                Model2 = c(trn_ml2, tst_ml2),
                Model3 = c(trn_ml3, tst_ml3),
                Model4 = c(trn_ml4, tst_ml4),
                Model5 = c(trn_ml5, tst_ml5))
```


```{r}
library(knitr)
kable(t(df), col.names = c("Training", "Testing"))
```

Among the five models, therefore, model3 is superior to the other models. Model3 uses displacement, weight, year, and origin as its predictors. 

***

## Exercise 5 (Simulating Multiple Regression)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \beta_3 x_{i3} + \beta_4 x_{i4} + \beta_5 x_{i5} + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = 2$
- $\beta_1 = -0.75$
- $\beta_2 = 1.5$
- $\beta_3 = 0$
- $\beta_4 = 0$
- $\beta_5 = 2$
- $\sigma^2 = 25$

We will use samples of size `n = 42`.

We will verify the distribution of $\hat{\beta}_2$ as well as investigate some hypothesis tests.

**(a)** We will first generate the $X$ matrix and data frame that will be used throughout the exercise. Create the following nine variables:

- `x0`: a vector of length `n` that contains all `1`
- `x1`: a vector of length `n` that is randomly drawn from a normal distribution with a mean of `0` and a standard deviation of `2`
- `x2`: a vector of length `n` that is randomly drawn from a uniform distribution between `0` and `4`
- `x3`: a vector of length `n` that is randomly drawn from a normal distribution with a mean of `0` and a standard deviation of `1`
- `x4`: a vector of length `n` that is randomly drawn from a uniform distribution between `-2` and `2`
- `x5`: a vector of length `n` that is randomly drawn from a normal distribution with a mean of `0` and a standard deviation of `2`
- `X`: a matrix that contains `x0`, `x1`, `x2`, `x3`, `x4`, and `x5` as its columns
- `C`: the $C$ matrix that is defined as $(X^\top X)^{-1}$
- `y`: a vector of length `n` that contains all `0`
- `sim_data`: a data frame that stores `y` and the **five** *predictor* variables. `y` is currently a placeholder that we will update during the simulation.

Report the sum of the diagonal of `C` as well as the 5th row of `sim_data`. For this exercise we will use the seed `420`. Generate the above variables in the order listed after running the code below to set a seed.

```{r}
set.seed(420)
sample_size = 42

x0 = rep(1, n)
x1 = rnorm(n, 0, 2)
x2 = runif(n, 0, 4)
x3 = rnorm(n, 0, 1)
x4 = runif(n, -2, 2)
x5 = rnorm(n, 0, 2)
X = cbind(x0, x1, x2, x3, x4, x5)
C = solve(t(X) %*% X)
y = rep(0, n)
sim_data = data.frame(y = y,
                      x1 = x1,
                      x2 = x2,
                      x3 = x3,
                      x4 = x4,
                      x5 = x5)
```

```{r}
# Sum of diagonal
sum(C[1,1], C[2,2], C[3,3], C[4,4], C[5,5], C[6,6])
```
```{r}
# 5th row of sim_data
sim_data[5,]
```


**(b)** Create three vectors of length `2500` that will store results from the simulation in part **(c)**. Call them `beta_hat_1`, `beta_3_pval`, and `beta_5_pval`.

```{r}
beta_hat_1 = rep(0, 2500)
beta_3_pval = rep(0, 2500)
beta_5_pval = rep(0, 2500)
```


**(c)** Simulate 2500 samples of size `n = 42` from the model above. Each time update the `y` value of `sim_data`. Then use `lm()` to fit a multiple regression model. Each time store:

- The value of $\hat{\beta}_1$ in `beta_hat_1`
- The p-value for the two-sided test of $\beta_3 = 0$ in `beta_3_pval`
- The p-value for the two-sided test of $\beta_5 = 0$ in `beta_5_pval`


```{r}
beta_0 = 2
beta_1 = -0.75
beta_2 = 1.5
beta_3 = 0
beta_4 = 0
beta_5 = 2
beta = rbind(beta_0, beta_1, beta_2, beta_3, beta_4, beta_5)
```

```{r}
for (i in 1:2500) {
  epsilon = rnorm(n, 0, 5)
  sim_data$y = beta_0 + beta_1*x1+ beta_2*x2 + beta_3*x3 + beta_4*x4 + beta_5*x5 + epsilon
  model = lm(y~., data=sim_data)

  beta_hat_1[i] = coef(model)['x1']
  beta_3_pval[i] = summary(model)$coefficients['x3','Pr(>|t|)']
  beta_5_pval[i] = summary(model)$coefficients['x5','Pr(>|t|)']
}
```


**(d)** Based on the known values of $X$, what is the true distribution of $\hat{\beta}_1$?

```{r}
sd1 = 5 * sqrt(C['x1','x1'])
sd1
```


Therefore, we expect $\hat{\beta_1}$to follow the following normal distribution.

$$\hat{\beta_1} \sim N(\beta_1=`r beta_1`, \sigma^2 = `r sd1`^2)$$
```{r}
beta_1
```


**(e)** Calculate the mean and variance of `beta_hat_1`. Are they close to what we would expect? Plot a histogram of `beta_hat_1`. Add a curve for the true distribution of $\hat{\beta}_1$. Does the curve seem to match the histogram?

```{r}
# Mean of beta_hat_1
(mean(beta_hat_1))

# Variance of beta_hat_1
(var(beta_hat_1))

# Standard deviation of beta_hat_1
(sd(beta_hat_1))
```


```{r}
hist(beta_hat_1, prob = TRUE, breaks = 25, border = "dodgerblue", 
            xlab = expression(hat(beta[1])))
curve(dnorm(x, mean = beta_1, sd = sd1),
      col = "darkorange", add = TRUE, lwd = 3)

```


**(f)** What proportion of the p-values stored in `beta_3_pval` is less than 0.10? Is this what you would expect?

```{r}
count = 0
for (i in 1:length(beta_3_pval)){
  if (beta_3_pval[i] < 0.1){
    count = count + 1
  }
}
# Expected P-value 
exp = coef(summary(model))['x3', 'Pr(>|t|)']

# Proportion
(ans = count / length(beta_3_pval))
```

There are only `r ans` which is less than 0.10. Considering my expected p-value is `r exp`, only few of the p-value turned out to be less than 0.1. Therefore, this result is what I have expected.  

**(g)** What proportion of the p-values stored in `beta_5_pval` is less than 0.01? Is this what you would expect?


```{r}
count = 0
for (i in 1:length(beta_5_pval)){
  if (beta_5_pval[i] < 0.01){
    count = count + 1
  }
}
# Expected P-value
exp = coef(summary(model))['x5', 'Pr(>|t|)']

# Proportion
(ans = count / length(beta_5_pval))
```

Considering my expected p-value is extremely small as in  `r exp`, it is expected that values in beta_5_pval should be less than 0.01. Therefore, the result above is an expected outcome.
